Complex Analysis of Askaryan Radiation: UHE-$\nu$ Identification and Reconstruction via the Hilbert Envelope of Observed Signals

### 2025-08-14 template.tex
updated first four plots; trying to gauge efficiency

\begin{abstract}
The detection of ultra-high energy neutrinos (UHE-$\nu$), with enegies above 10 PeV, has been a long-time goal of the astroparticle physics community.  A key strategy has been to deploy autonomous, radio-frequency (RF) detetectors in polar regions that rely on the Askaryan effect in ice for the neutrino signal.  The Askaryan effect occurs when the excess negative charge within a high-energy cascade radiates in a dense medium.  UHE-$\nu$ can induce cascades that radiate in the radio-frequency (RF) bandwidth above thermal backgrounds.  To identify UHE-$\nu$ signals in future data from Askaryan-class detectors, analytic models of the Askaryan electromagnetic field have been created and compared to simulations and laboratory measurements.  In the past, these models have only described the Askaryan electromagnetic field, leaving the effect of the RF detection channels on the field to simulation packages.  In this work, we present a fully analytic Askaryan model that accounts for the effect of an RF detection channel.  We provide formulas for the observed voltage trace, and for the Hilbert envelope of the trace.  We match the equations to data computed with NuRadioMC, a key Monte Carlo toolset in UHE-$\nu$ detection.  Our analytic model and NuRadioMC match, with correlation coefficients $>0.95$.  Using this correlation, we demonstrate that RF thermal noise and UHE-$\nu$ signals can be classified to better than xxx.  Finally, we reconstruct the logarithm of the UHE-$\nu$ cascade energy using the analytic model, with a precision of yyy.
\end{abstract}

>>>>

\begin{abstract}
The detection of ultra-high energy neutrinos (UHE-$\nu$), with enegies above 10 PeV, has been a long-time goal of the astroparticle physics community.  A key strategy has been to deploy autonomous, radio-frequency (RF) detetectors in polar regions that rely on the Askaryan effect in ice for the neutrino signal.  The Askaryan effect occurs when the excess negative charge within a high-energy cascade radiates in a dense medium, and UHE-$\nu$ can induce such cascades that radiate in the radio-frequency (RF) bandwidth above thermal backgrounds.  To identify UHE-$\nu$ signals in future data from Askaryan-class detectors, analytic models of the Askaryan electromagnetic field have been created and compared to simulations and laboratory measurements.  These models have correctly described the Askaryan electromagnetic field, while leaving the effect of the RF detection channels on the field to simulation packages.  In this work, we present a fully analytic Askaryan model that accounts for the effect of an RF detection channel.  We provide formulas for the observed voltage trace, and for the Hilbert envelope of the trace.  We match the equations to data computed with NuRadioMC, a key Monte Carlo toolset in UHE-$\nu$ detection.  Our analytic model and NuRadioMC match, with correlation coefficients $>0.95$.  Using this correlation, UHE-$\nu$ signals can be distinguished from RF backgrounds with an efficiency of xxx.  Finally, we reconstruct the logarithm of the UHE-$\nu$ cascade energy with a precision of yyy using a single string of RF channels.
\end{abstract}
